---
title: "Th Stopping in Philadelphia Puerto Rican English"
output:
  html_notebook:
    toc: yes
    toc_float: yes
  html_document:
    toc: yes
    df_print: paged
---
# Setup
## Set random seed

This ensures that all results are replicable
```{r}
set.seed(5241988)
```

## Load Packages

Define a function that will take a list of required packages, install any that are missing, and add all packages to the local workspace:

```{r}
get_packages<-function(package_list){
  # Install packages not yet installed
  installed_packages <- package_list %in% rownames(installed.packages())
  if (any(installed_packages == FALSE)) {
    install.packages(package_list[!installed_packages])
  }
  # Packages loading
  invisible(lapply(package_list, require, character.only = TRUE))}
```

Apply to the package list for this analysis:

```{r, print=FALSE,warning=FALSE,results=FALSE}
packages<-c('dplyr','tidyr','magrittr','ggplot2','statcomp','bayesplot','rstanarm','bayestestR','BayesFactor','irr','gridExtra','parallel','shinystan')
get_packages(packages)
```

Set some environmental variables:

```{r, results=FALSE}
options(mc.cores=detectCores())
test_regex="^scale.*|^\\(Int.*"
regex="^\\(Int|^Age.*|^Sex.*|^Birth.*|^s\\(B.*"
my_probs=c(0.0275,0.5,0.975)
```

## Set up Workspace

### Set Session Variables

Set color scheme and line types for tokens

```{r, results=FALSE}
voiced_col<-RColorBrewer::brewer.pal(4, "Paired")[1]
voiced_col_alt<-RColorBrewer::brewer.pal(4, "Paired")[2]
voiceless_col<-RColorBrewer::brewer.pal(4, "Paired")[3]
voiceless_col_alt<-RColorBrewer::brewer.pal(4, "Paired")[4]
male_col<-RColorBrewer::brewer.pal(4, "Accent")[2]
female_col<-RColorBrewer::brewer.pal(4, "Accent")[3]
female_line<-'solid'
male_line<-'dashed'
```

### Set Plot Theme

Set general theme for plots

```{r, results=FALSE}
theme_set(theme_minimal()+
            theme(plot.title = element_text(size=22),
                  legend.key.size = unit(1,"cm"),
                  legend.text = element_text(size=12),
                  legend.title = element_text(size=15),
                  panel.spacing=unit(2,"lines"),
                  panel.background=element_rect(color='black'),
                  strip.text=element_text(color='black',size=18),
                  axis.text=element_text(color='black',size=12),
                  axis.title=element_text(color='black',size=15)))
```

### Pre-load models
By default, the chunks in this notebook that pertain to the statistical analysis have been set not to evaluate. This will load all models to the workspace so they can be queried without having to wait for them to run.
```{r}
load('../Data/all_models_final.RData')
```

# Test Set Data

By default, we'll assume that the script hasn't been moved from its original location and that all data files are in the same directory as this script. If that's not true, use `setwd()` to change the working directory first.

## Data Processing

Load in test set and process

```{r}
test_set=read.delim('../Data/testset_results.txt')
process_test_set=function(data){
  data$Speaker=with(data,factor(ifelse(substr(File,1,1)=='A',"A","G")))
  data$File=as.factor(data$File)
  data$Phone=as.factor(data$Phone)
  data$Word=as.factor(data$Word)
  data$Mean_HNR=as.numeric(data$Mean_HNR)
  return(data)}
test_set %<>% process_test_set()
```

Divide into voiceless and voiced

```{r}
voiced_test_set<-test_set %>% filter(.,Phone=='DH'|Phone=='D') %>% droplevels()
voiceless_test_set<-test_set %>% filter(.,Phone=="TH"|Phone=="T") %>% droplevels()
voiced_test_set$Phone<-relevel(voiced_test_set$Phone,ref="DH")
voiceless_test_set$Phone<-relevel(voiceless_test_set$Phone,ref="TH")
```

## Data Visualization

```{r}
phones<-rev(c('/θ/','/t/','/ð/','/d/'))
cog<-ggplot(test_set,aes(y=Phone,x=CoG,fill=Phone))+geom_violin()+stat_summary(geom='errorbar',width=0.25,fun.data='mean_cl_boot')+scale_fill_brewer(palette='Paired')+scale_y_discrete(labels=phones)+coord_cartesian(xlim=c(0,1500))+theme(legend.position='none')+labs(x="Center of Gravity")
skew<-ggplot(test_set,aes(x=Skewness,y=Phone,fill=Phone))+geom_violin()+stat_summary(geom='errorbar',width=0.25,fun.data='mean_cl_boot')+scale_fill_brewer(palette='Paired')+scale_y_discrete(labels=phones)+coord_cartesian(xlim=c(0,25))+theme(legend.position='none')
kur<-ggplot(test_set,aes(x=Kurtosis,y=Phone,fill=Phone))+geom_violin()+stat_summary(geom='errorbar',width=0.25,fun.data='mean_cl_boot')+scale_fill_brewer(palette='Paired')+scale_y_discrete(labels=phones)+coord_cartesian(xlim=c(0,500))+theme(legend.position='none')
hnr<-ggplot(test_set,aes(x=Mean_HNR,y=Phone,fill=Phone))+geom_violin()+stat_summary(geom='errorbar',width=0.25,fun.data='mean_cl_boot')+scale_fill_brewer(palette='Paired')+scale_y_discrete(labels=phones)+theme(legend.position='none')+labs(x="Mean HNR")
grid.arrange(cog,hnr,skew,kur)->figure1
```

## Data Analysis

### Run models

#### Voiceless Data

```{r, results='hide',eval=F}
ms_testset=stan_glmer(Phone~scale(Mean_HNR)+scale(CoG)+scale(Skewness)+scale(Kurtosis)+(1|Speaker),voiceless_test_set,family='binomial',adapt_delta=0.99,warmup=2000,iter=6000,chains=8)
ms_testset_loo<-loo(ms_testset,k_threshold=0.7)
```

#### Voiced Data

```{r, results='hide',eval=F}
mv_testset=stan_glmer(Phone~scale(Mean_HNR)+scale(CoG)+scale(Skewness)+scale(Kurtosis)+(1|Speaker),voiced_test_set,family='binomial',adapt_delta=0.99,warmup=2000,iter=6000,chains=8)
mv_testset_loo<-loo(mv_testset,k_threshold=0.7)
```

### View Results

#### Model Summary

```{r}
summary(ms_testset,regex_pars=test_regex,probs=my_probs,digits=3)
```

```{r}
summary(mv_testset,regex_pars=test_regex,probs=my_probs,digits=3)
```
#### HDIs and Bayes Factors

```{r}
print("Voiceless Data:")
hdi(ms_testset)
bayesfactor(ms_testset)
```
```{r}
print("Voiced Data:")
hdi(mv_testset)
bayesfactor(mv_testset)
```



#### Posterior Draws

```{r}
color_scheme_set(c(voiceless_col,rep("#000000",5)))
ms_testset_post<-mcmc_areas(ms_testset,regex_pars=test_regex,prob_outer=0.95)+ggtitle("Voiceless")+scale_y_discrete(labels=c("Intercept","Mean HNR","CoG","Skewness","Kurtosis"))+geom_vline(xintercept=0,linetype='dotted')+labs(x="Stop Likelihood")
color_scheme_set(c(voiced_col,rep("#000000",5)))
mv_testset_post<-mcmc_areas(mv_testset,regex_pars=test_regex,prob_outer=0.95)+ggtitle("Voiced")+scale_y_discrete(labels=c("Intercept","Mean HNR","CoG","Skewness","Kurtosis"))+geom_vline(xintercept=0,linetype='dotted')+labs(x="Stop Likelihood")
grid.arrange(ms_testset_post,mv_testset_post,ncol=2)->figure2
```

#### Region of Practical Equivalence (ROPE)

```{r,warning=FALSE}
ms_testset_rope<-plot(rope(ms_testset,range=rope_range(ms_testset),ci_method='hdi'),alpha=0.75)+ggtitle("Voiceless")+scale_fill_manual(values=c(voiceless_col,voiceless_col_alt))
mv_testset_rope<-plot(rope(mv_testset,range=rope_range(mv_testset),ci_method='hdi'),alpha=0.75)+ggtitle("Voiced")+scale_fill_manual(values=c(voiced_col,voiced_col_alt))
grid.arrange(ms_testset_rope,mv_testset_rope,ncol=1)
```

#### Shinystan App (will open in new browser if code is run)

```{r, eval=FALSE}
launch_shinystan(ms_testset)
```

```{r,eval=FALSE}
launch_shinystan(mv_testset)
```

# PREP Data

PREP = Puerto Rican English in Philadelphia Corpus

[More info](https://osf.io/7km4r/)

## Data Processing

### Load Data from Sources

By default, we'll assume that the script hasn't been moved from its original location and that all data files are in the same directory as this script. If that's not true, use `setwd()` to change the working directory first.

```{r}
prep_data=read.delim('../Data/prep_data.txt')
socio_data=read.csv('../Data/prep_socio.csv')
perceptual_data=read.delim('../Data/perceptual_data.txt')
perceptual_tokens_voicing_codes=read.delim('../Data/voicing_codes.txt')
```

### Process Data

Start by getting phonetic context

```{r}
prep_data$FollowingStress=as.factor(with(prep_data,substr(FollowingPhone,nchar(FollowingPhone),nchar(FollowingPhone))))
prep_data$FollowingVowel=as.factor(with(prep_data,substr(FollowingPhone,1,nchar(FollowingPhone)-1)))
prep_data$FollowingPhone=as.factor(prep_data$FollowingPhone)
```

Remove tokens that aren't prevocalic (i.e., are consonant clusters)

```{r}
prep_data %<>% filter(!(FollowingPhone %in% list('R','W','Y'))) 
prep_data %<>% mutate(Voicing=if_else(Phone %in% list('D','DH'),'Voiced','Voiceless'),Manner=if_else(Phone %in% list('D','T'),'Stop','Fricative'),Metric=if_else(Voicing=='Voiced','Mean_HNR','Skewness'))
```

Join to sociodemographic data and specify data structure

```{r,warning=FALSE}
prep_data %>% left_join(.,socio_data, by=c('File'='Speaker'))->prep_socio # Merge socio data with speaker data
prep_socio$Sex=as.factor(prep_socio$Sex)
prep_socio$Mean_HNR=as.numeric(prep_socio$Mean_HNR)
prep_socio$File=as.factor(prep_socio$File)
prep_socio$Word=gsub("\\(\\(","",prep_socio$Word)
prep_socio$Word=gsub("\\)\\)","",prep_socio$Word)
prep_socio$Word=as.factor(prep_socio$Word)
prep_socio$Phone=as.factor(prep_socio$Phone)
prep_socio %<>% mutate(AgeGroup=if_else(BirthYear>=1985,'Younger','Older'),Speaker=File)
```

Get average values by speaker, voicing, and phonetic context

```{r,warning=FALSE}
prep_data$PrecedingPhone=as.factor(prep_data$PrecedingPhone)
prep_data$PrecedingWord=as.factor(prep_data$PrecedingWord)
prep_data$Skewness=as.numeric(prep_data$Skewness)
prep_data$Mean_HNR=as.numeric(prep_data$Mean_HNR)

speaker_avgs<-prep_data %>% filter(Phone=='D' | Phone=='T') %>% 
  group_by(File, Phone, FollowingVowel, FollowingStress) %>% 
  reframe(count=n(),HNR_stop=mean(na.omit(Mean_HNR)),sd_HNR_stop=sd(na.omit(Mean_HNR)), skewness_stop=mean(na.omit(Skewness)),sd_skewness_stop=sd(na.omit(Skewness))) %>%
  ungroup()

voiced_avgs<-speaker_avgs %>% filter(Phone=='D') %>% select(-Phone)
voiceless_avgs<-speaker_avgs %>% filter(Phone=='T') %>% select(-Phone)
```

Functions to calculate normalized stopping ratios and remove outliers \>3 standard deviations from the group mean, by speaker.

```{r}
make_stopping_ratios<-function(data,p){
  data %<>% filter(Phone==p) %>% 
  left_join(voiced_avgs,by=c('File','FollowingVowel','FollowingStress')) %>% 
  mutate(HNR_ratio=Mean_HNR/HNR_stop,skewness_ratio=Skewness/skewness_stop) %>%
  mutate(ratio=(if_else(Phone=='TH',skewness_ratio,if_else(Phone=="DH",HNR_ratio,NA)))) %>%
  left_join(socio_data,by=c('File'='Speaker')) %>% rename(Speaker=File) %>%
  mutate(AgeGroup=as.factor(ifelse(BirthYear>=1985,'Younger','Older')))
  return(data)
}

remove_outliers<-function(data){
  data %>% filter(!is.na(ratio)) %>% mutate(Pseudonym=ifelse(.$Speaker=='S39','Carina',.$Pseudonym))%>%
    mutate(mean_ratio_byspeaker=mean(ratio),sd_ratio_byspeaker=sd(ratio)) %>%
    ungroup() %>%
    filter(abs(ratio-mean_ratio_byspeaker)<=3*sd_ratio_byspeaker) %>%
    filter(abs(ratio-mean(ratio))<=3*sd(ratio))->result
  print(paste0("Removed ",nrow(data)-nrow(result)," rows, or ",round(100*(1-nrow(result)/nrow(data)),2),"% of data."))
  return(result)
}
```

Apply to data and create production dataset

```{r,warning=FALSE}
voiceless<-make_stopping_ratios(prep_data,"TH") %>% remove_outliers()
voiced<-make_stopping_ratios(prep_data,"DH") %>% remove_outliers()
production_data<-rbind(voiced,voiceless)
```

## Data Visualization

### Data distribution

```{r, echo=T,fig.keep='all'}
production_data %>% group_by(Word,Phone) %>% summarise(count=n()) %>% arrange(desc(count)) %>%  {.->> word_cts} %>%
  ggplot(.,aes(x=count))+geom_density()+scale_x_continuous(breaks=seq(0,500,by=50))+coord_cartesian(xlim=c(0,500))+ggtitle("Word Distribution")

```

Note that only 2 of the top 10 words are /θ/ words (think,things)

```{r}
word_cts %>%
  filter(count>=40&(Phone=='DH'|Phone=='TH')) %>% 
  arrange(desc(count)) %>% 
  head(20) 
```

Top 10 /ð/ words
```{r}
word_cts %>%
  filter(Phone=='DH') %>% 
  arrange(desc(count)) %>% 
  head(10) 
```

Top 10 /θ/ words
```{r}
word_cts %>%
  filter(Phone=='TH') %>% 
  arrange(desc(count)) %>% 
  head(10) 
```


Visualize distribution of fricatives and stops across data by metric

```{r,fig.keep='all'}

ggplot(prep_socio,aes(x=if_else(Voicing=='Voiced',Mean_HNR,Skewness),fill=Manner))+
  geom_density(alpha=0.5)+facet_grid(Voicing~.)+
  labs(x="Metric",y="Density")+scale_fill_viridis_d(option='cividis',begin=0.3,end=0.9)+theme(legend.position='bottom')

ggplot(prep_socio,aes(x=if_else(Voicing=='Voiced',Mean_HNR,Skewness),fill=Manner))+
  geom_density(alpha=0.5)+facet_grid(Voicing~Sex)+
  labs(x="Metric",y="Density")+scale_fill_viridis_d(option='viridis',begin=0.2,end=0.6)+theme(legend.position='bottom')
 
ggplot(prep_socio,aes(x=if_else(Voicing=='Voiced',Mean_HNR,Skewness),fill=Manner))+
  geom_density(alpha=0.5)+facet_grid(Sex~AgeGroup+Voicing)+
  labs(x="Metric",y="Density")+scale_fill_viridis_d(option='inferno')+theme(legend.position='bottom')

```

Individual ranges to see if there are any major outliers that remain after trimming procedure

```{r,fig.keep='all'}
ggplot(voiceless,aes(x=Pseudonym,y=ratio))+
  geom_boxplot(fill=voiceless_col)+
  labs(x="Speaker",y="Normalized Stopping Ratio\n(Skewness)")+
  ggtitle("Voiceless")+scale_y_continuous(breaks=seq(-1,5,by=0.5))+geom_hline(yintercept=1,linetype='dashed')+
  coord_flip()->p_voiceless_byspeaker
ggplot(voiced,aes(x=Pseudonym,y=ratio))+
  geom_boxplot(fill=voiced_col)+
  labs(x="Speaker",y="Normalized Stopping Ratio\n(Mean HNR)")+
  scale_y_continuous(breaks=seq(-1,5,by=1))+
  ggtitle("Voiced")+geom_hline(yintercept=1,linetype='dashed')+
  coord_flip()->p_voiced_byspeaker
grid.arrange(p_voiceless_byspeaker,p_voiced_byspeaker,ncol=2)
```

### Basic Sociolinguistic Trends

View general trends by birth year

```{r}
age_voiceless<-ggplot(voiceless,aes(x=BirthYear,y=skewness_ratio-1,linetype=Sex))+
  geom_hline(yintercept=0,linewidth=0.5,linetype='dotted')+geom_rug(sides="br",linetype='solid')+
  geom_smooth(aes(group=Sex),color='black',fill=voiceless_col,method='gam',formula=y~s(x,k=3,m=2))+
  scale_fill_manual(values=c(male_col,female_col))+
  scale_y_continuous(breaks=seq(-1.5,1.5,by=0.5))+geom_point(alpha=0.05,position=position_jitter(0.3))+
  labs(x="Birth Year",y="Skewness Ratio\n(Deviance from 1)")+coord_cartesian(ylim=c(-1.5,1.5))+
  ggtitle("Voiceless")+theme(legend.position='bottom',strip.text=element_text(size=15))+scale_x_continuous(expand = expansion(mult = 0.1))

age_voiced<-ggplot(voiced,aes(x=BirthYear,y=HNR_ratio-1,linetype=Sex))+
  geom_hline(yintercept=0,linewidth=0.5,linetype='dotted')+geom_rug(sides="br",linetype='solid')+
  geom_smooth(aes(group=Sex),color='black',fill=voiced_col,method='gam',formula=y~s(x,k=3,m=2))+
  scale_y_continuous(breaks=seq(-1.5,1.5,by=0.5))+geom_point(alpha=0.05,position=position_jitter(0.3))+
  labs(x="Birth Year",y="HNR Ratio\n(Deviance from 1)")+
  ggtitle("Voiced")+coord_cartesian(ylim=c(-1.5,1.5))+
  theme(legend.position='bottom',strip.text=element_text(size=15))+scale_x_continuous(expand = expansion(mult = 0.1))

grid.arrange(age_voiceless,age_voiced,ncol=2)->figure4
print(figure4)
```

View trends by birth year, grouping by speaker

```{r}
age_voiced_byspeaker<-ggplot(voiced,aes(x=BirthYear,y=HNR_ratio))+
  geom_hline(yintercept=1,linewidth=0.5,linetype='dotted')+
  geom_boxplot(aes(group=Speaker),fill=voiced_col)+
  labs(x="Birth Year",y="Normalized HNR Ratio")+ggtitle("Voiced")+
  theme(legend.position='bottom')+scale_x_continuous(expand = expansion(mult = 0.1))+
  facet_wrap(~Sex,ncol=1)

age_voiceless_byspeaker<-ggplot(voiceless,aes(x=BirthYear,y=skewness_ratio))+
  geom_hline(yintercept=1,linewidth=0.5,linetype='dotted')+
  geom_boxplot(aes(group=Speaker),fill=voiceless_col)+
  labs(x="Birth Year",y="Normalized Skewness Ratio")+ggtitle("Voiceless")+
  theme(legend.position='bottom')+
  scale_x_continuous(expand = expansion(mult = 0.1))+
  facet_wrap(~Sex,ncol=1)

grid.arrange(age_voiceless_byspeaker,age_voiced_byspeaker,ncol=2)
```

## Data Analysis

Set Contrasts

```{r}
set_contrasts<-function(data){
  data$Sex<-as.factor(data$Sex)
  data$Sex<-relevel(data$Sex,ref='Male')
  data$AgeGroup=as.factor(data$AgeGroup)
  data$AgeGroup<-relevel(data$AgeGroup,ref="Older")
  return(data)
  }
```

```{r}
voiced %<>% set_contrasts()
voiceless %<>% set_contrasts()
voiced$HNR_ratio_deviance=voiced$HNR_ratio-1
voiceless$skewness_ratio_deviance=voiceless$skewness_ratio-1
```

### Run Models

#### Voiceless Data
```{r,eval=F}
ms_prep_gamm<-stan_gamm4(skewness_ratio_deviance~Sex*BirthYear+s(BirthYear,bs='tp',by=Sex,k=3,m=2),random=~(1|Speaker),data=voiceless,adapt_delta=0.999,warmup=2000,iter=6000,chains=8)
ms_prep_gamm_loo<-loo(ms_prep_gamm,k_threshold=0.7)
```


#### Voiced Data
```{r,eval=F}
mv_prep_gamm<-stan_gamm4(HNR_ratio_deviance~Sex*BirthYear+s(BirthYear,bs='tp',by=Sex,k=3,m=2),random=~(1|Speaker),data=voiced,adapt_delta=0.999,warmup=2000,iter=6000,chains=8)
mv_prep_gamm_loo<-loo(mv_prep_gamm,k_threshold=0.7)
```

### View Results


#### Model Summary

##### Voiceless

```{r}
summary(ms_prep_gamm,regex_pars=regex,probs=my_probs,digits=3)
```

##### Voiced

```{r}
summary(mv_prep_gamm,regex_pars=regex,probs=my_probs,digits=3)

```

##### HDIs and Bayes Factors

##### Voiceless
```{r}
hdi(ms_prep_gamm)
bayesfactor(ms_prep_gamm)
```
##### Voiced
```{r}
hdi(mv_prep_gamm)
bayesfactor(mv_prep_gamm)
```

##### Posterior Draws

```{r,fig.keep='all'}
color_scheme_set(c(voiceless_col,rep("#000000",5)))
ms_prep_gamm_post<-mcmc_areas(ms_prep_gamm,regex_pars=regex,prob_outer=0.95)+ggtitle("Voiceless")
color_scheme_set(c(voiced_col,rep("#000000",5)))
mv_prep_gamm_post<-mcmc_areas(mv_prep_gamm,regex_pars=regex,prob_outer=0.95)+ggtitle("Voiced")
grid.arrange(ms_prep_gamm_post,mv_prep_gamm_post,ncol=2)
```

Nonlinear Plot

```{r fig.keep='all'}
color_scheme_set(c(voiceless_col,rep("#000000",5)))
ms_prep_gamm_nl<-plot_nonlinear(ms_prep_gamm,prob=0.95)+ggtitle("Voiceless Data")
color_scheme_set(c(voiced_col,rep("#000000",5)))
mv_prep_gamm_nl<-plot_nonlinear(mv_prep_gamm,prob=0.95)+ggtitle("Voiced Data")
grid.arrange(ms_prep_gamm_nl,mv_prep_gamm_nl,ncol=1)
```

##### Region of Practical Equivalence (ROPE)
##### Voiceless
```{r}
ms_prep_gamm_rope<-plot(rope(ms_prep_gamm,range=rope_range(ms_prep_gamm),ci_method='hdi'))+ggtitle("Voiceless")+scale_fill_manual(values=c(voiceless_col,voiceless_col_alt))+theme(legend.position='bottom')
ms_prep_gamm_rope
```
##### Voiced
```{r}
mv_prep_gamm_rope<-plot(rope(mv_prep_gamm,range=rope_range(mv_prep_gamm),ci_method='hdi'))+ggtitle("Voiced")+scale_fill_manual(values=c(voiced_col,voiced_col_alt))+theme(legend.position='bottom')

mv_prep_gamm_rope
```

##### Shinystan Apps (will open in new browser if code is run)

Voiceless Data

```{r, eval=FALSE}
launch_shinystan(ms_prep_gamm)
```

Voiced Data

```{r, eval=FALSE}
launch_shinystan(mv_prep_gamm)
```

# Perceptual Coding

## Data Processing

```{r}
perceptual_data %>% filter(Coder!="") %>%
  mutate(token=paste(Speaker,Word,Start,End,sep="_")) %>% 
  left_join(.,perceptual_tokens_voicing_codes,by='Word') %>% 
  left_join(.,socio_data,by='Speaker') %>% 
  mutate(AgeGroup=ifelse(as.numeric(BirthYear)>=1985,'Younger','Older')) %>%
  filter(Code!="Unsure")->perceptual_dataset_raw
perceptual_dataset_raw %<>% set_contrasts()
```

```{r}
perceptual_dataset_raw %>% 
  group_by(Speaker,AgeGroup,BirthYear,Sex,Voicing,Code) %>% 
  tally() %>% pivot_wider(names_from=Code,values_from=n) %>% 
  mutate(Total=Stop+Fricative,Stopping_Rate=100*Stop/(Total)) %>% 
  {.->> perceptual_rates_raw} %>% print()
```

### Labov (dh) Index Comparison
```{r}
perceptual_data %>% filter(Coder!="") %>%
  mutate(token=paste(Speaker,Word,Start,End,sep="_")) %>% 
  left_join(.,perceptual_tokens_voicing_codes,by='Word') %>% 
  left_join(.,socio_data,by='Speaker') %>% 
  mutate(AgeGroup=ifelse(as.numeric(BirthYear)>=1985,'Younger','Older'),dh_value=case_match(Code,"Unsure"~1,"Fricative"~0,"Stop"~2)) %>% {.->>perceptual_group_level} %>%
  group_by(AgeGroup,Sex) %>% reframe(dh_index=100*mean(dh_value))

perceptual_group_level %>% 
  group_by(Voicing,AgeGroup,Sex,Code) %>% 
  filter(Code!="Unsure") %>%
  tally() %>% group_by(AgeGroup,Sex,Voicing) %>% 
  pivot_wider(names_from=Code,values_from=n) %>% 
  mutate(Total=Stop+Fricative,Stopping_Rate=100*Stop/(Total))

perceptual_group_level %>% filter(Coder!="") %>% group_by(Sex,AgeGroup) %>% reframe(count=n_distinct(Speaker))
```

### Rater Consensus

Get Fleiss' Kappa for IRR

```{r}
perceptual_data %>% filter(Coder!="") %>% 
  mutate(token=paste(Speaker,Word,Start,End,sep="_")) %>% 
  select(token,Coder,Code) %>% 
  pivot_wider(names_from=`Coder`,values_from=Code) %>% {. ->> perceptual_irr} %>%
  kappam.fleiss()
```

Show tokens by level of agreement
```{r}
perceptual_irr %>% rowwise %>%
        mutate(distinct = n_distinct(unlist(across(starts_with("A"))))) %>%
        ungroup() -> perceptual_irr_agreement

perceptual_irr_agreement
```
## Data Visualization

Stopping Rates by Sex and Birth Year based on Individual Rater Score

```{r}
perceptual_voiceless_rates_raw<- perceptual_rates_raw %>%
  filter(Voicing=="Voiceless") %>%
  ggplot(aes(x=BirthYear,y=Stopping_Rate,linetype=Sex))+
  geom_rug(sides="br",linetype='solid')+
  geom_smooth(aes(group=Sex),color='black',fill=voiceless_col,method='gam',formula=y~s(x,k=3,m=2))+
  #scale_y_continuous(breaks=seq(-1.5,1.5,by=0.5))+
  geom_point(alpha=0.5,position=position_jitter(0.3))+
  labs(x="Birth Year",y="Stopping Rate\n(Perceptual, by Rater)")+
  ggtitle("Voiceless")+scale_linetype_manual(values=c(male_line,female_line))+
  theme(legend.position='bottom',strip.text=element_text(size=15))+scale_x_continuous(expand = expansion(mult = 0.1))

perceptual_voiced_rates_raw<- perceptual_rates_raw %>%
  filter(Voicing=="Voiced") %>%
  ggplot(aes(x=BirthYear,y=Stopping_Rate,linetype=Sex))+
  geom_rug(sides="br",linetype='solid')+
  geom_smooth(aes(group=Sex),color='black',fill=voiced_col,method='gam',formula=y~s(x,k=3,m=2))+
  geom_point(alpha=0.5,position=position_jitter(0.3,0.3))+
  labs(x="Birth Year",y="Stopping Rate\n(Perceptual, by Rater)")+
  ggtitle("Voiced")+scale_linetype_manual(values=c(male_line,female_line))+
  theme(legend.position='bottom',strip.text=element_text(size=15))+scale_x_continuous(expand = expansion(mult = 0.1))

grid.arrange(perceptual_voiceless_rates_raw,perceptual_voiced_rates_raw,ncol=2)
```


## Data Analysis

### Run Models

Force outcome variable to be factor for binomial gam

```{r}
perceptual_dataset_raw$Code %<>% as.factor()
```

#### Voiceless Data

```{r,eval=F}
ms_per_raw_gamm<-stan_gamm4(Code~BirthYear*Sex+s(BirthYear,by=Sex,k=3,m=2,bs='tp'),random=~(1|Speaker),data=subset(perceptual_dataset_raw,Voicing=="Voiceless"),family='binomial',adapt_delta=0.999,warmup=2000,iter=6000,chains=8)

ms_per_raw_gamm_loo<-loo(ms_per_raw_gamm)
```

#### Voiced Data

```{r,eval=FALSE}
mv_per_raw_gamm<-stan_gamm4(Code~BirthYear*Sex+s(BirthYear,by=Sex,k=3,m=2,bs='tp'),random=~(1|Speaker),data=subset(perceptual_dataset_raw,Voicing=="Voiced"),family='binomial',adapt_delta=0.999,warmup=2000,iter=6000,chains=8)
mv_per_raw_gamm_loo<-loo(mv_per_raw_gamm,k_threshold=0.7)
```


### View Results

#### Model Summary

##### Voiceless Data

```{r}
summary(ms_per_raw_gamm,probs=my_probs,regex_pars=regex,digits=3)
```

##### Voiced Data

```{r}
summary(mv_per_raw_gamm,probs=my_probs,regex_pars=regex,digits=3)
```

#### Nonlinear Plot

```{r fig.keep='all'}
color_scheme_set(c(voiceless_col,rep("#000000",5)))
ms_per_raw_gamm_nl<-plot_nonlinear(ms_per_raw_gamm,prob=0.95)+ggtitle("Voiceless Data")
color_scheme_set(c(voiced_col,rep("#000000",5)))
mv_per_raw_gamm_nl<-plot_nonlinear(mv_per_raw_gamm,prob=0.95)+ggtitle("Voiced Data")
grid.arrange(ms_per_raw_gamm_nl,mv_per_raw_gamm_nl,ncol=1)
```

#### HDI and Bayes Factors

##### Voiceless Data
```{r}
hdi(ms_per_raw_gamm)
bayesfactor(ms_per_raw_gamm)
rope(ms_per_raw_gamm,range=rope_range(ms_per_raw_gamm),ci_method='hdi',ci=0.95)
```

#### Voiced Data
```{r}
hdi(mv_per_raw_gamm)
bayesfactor(mv_per_raw_gamm)
rope(mv_per_raw_gamm,range=rope_range(mv_per_raw_gamm),ci_method='hdi',ci=0.95)
```
#### Posterior Draws

```{r}
color_scheme_set(c(voiceless_col,rep("#000000",5)))
ms_per_raw_post<-mcmc_areas(ms_per_raw,regex_pars=regex,prob_outer=0.95)+ggtitle("Voiceless")
color_scheme_set(c(voiced_col,rep("#000000",5)))
mv_per_raw_post<-mcmc_areas(mv_per_raw,regex_pars=regex,prob_outer=0.95)+ggtitle("Voiced")
grid.arrange(ms_per_raw_post,mv_per_raw_post,ncol=1)
```

#### Region of Practical Equivalence (ROPE)

```{r}
ms_per_raw_rope<-plot(rope(ms_per_raw,range=rope_range(ms_per_raw)))+ggtitle("Voiceless")+scale_fill_manual(values=c(voiceless_col,voiceless_col_alt))
mv_per_raw_rope<-plot(rope(mv_per_raw,range=rope_range(mv_per_raw)))+ggtitle("Voiced")+scale_fill_manual(values=c(voiced_col,voiced_col_alt))
grid.arrange(ms_per_raw_rope,mv_per_raw_rope,ncol=1)
```

##### Shinystan Apps (will open in new browser if code is run)

Voiceless Data

```{r,eval=F}
launch_shinystan(ms_per_raw)
```

Voiced Data

```{r,eval=F}
launch_shinystan(mv_per_raw)
```


```{r}
#save.image('../Data/all_models_final.RData')
```
